---
title: "hw3"
author: "Ritika Agrawal"
github username: "ritikagrawal1710"
date: "February 19, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Homework #3
## Ritika Agrawal ; ritikagrawal1710
## Tips at Restaurants

In this homework we will use the `tips` data set. This data set is part of the `reshape2` package. You can load the data set by executing the command:
```{r}
library(dplyr)
library(ggplot2)
data(tips, package="reshape2")
head(tips)
```
If you do not have available the package `reshape2`, issue `install.packages('reshape2')` to install it. The information contained in the data is collected by one waiter, who recorded over the course of a season information about each tip he received working in one restaurant. See `?tips` for a description of all of the variables. 

**Submission instructions**: Create a folder named `ds202_hw3`, and name the RMarkDown file `hw3.Rmd`  which should include your solutions under this folder. For submission, create a GitHub repository named `ds202_hw3` under your GitHub account, and push both `hw3.Rmd` and the knitted `hw3.html` before the deadline. I will assume you use the same GitHub username as for your HW2 submission. The latest version of your homework appearing on GitHub before the deadline is graded. *It is crucial to follow the exact folder structure and file names*, so that your homework can be reproduced and graded by an automated script. This homework is **due on Feb 19** before class.

1.  How many parties did the waiter serve? Store the number in `numParty` and print. 
```{r}
numParty = nrow(tips)
numParty
```

2. What are the types of variables in the data frame `tips`? Include the code and also explain verbally.
```{r}
head(tips)
summary(tips)
```
Response: 'tips' is dependent variable. The others are independent variables. The variables, total_bill, size are numerical and the variables sex, smoker,day,time are categorical variables

3. Create a vector named `day` containing the day of visits in `tips`. The factor levels should be ordered from Thursday to Sunday. Print the variable.
```{r}
my_group <- group_by(tips, day)
day=summarize(my_group, Value = n())
day$day <- factor(day$day, levels= c("Monday", 
                                         "Tuesday", "Wednesday", "Thur", "Fri", "Sat","Sun"))
day = day[order(day$day), ]
day
```

4. Create a data frame named `female5` containing the meal paid by a female payer in a party with size greater than or equal to 5. Print the data frame.
```{r}
female5 <- subset(tips, size >= 5 & sex == "Female")
head(female5)
```


5. How does the tipping amount (`tip`) depend on the overall bill (`total_bill`)? Use the `ggplot2` package to make a chart. Describe the relationship in words.  
```{r}
ggplot(tips, aes(x=total_bill, y=tip)) + geom_point()
```

Response: Relation seems to be linear

6. Describe at least two types of anomalies in the previous plot. What do they mean?

Response: If we see the data points 32,2 and 7,5, these are called outliers which do not follow the pattern as most other data points. 


7. Introduce a variable `tiprate` into the data set which stands for the rate of tips. What is the average rate for tips? Explain verbally. 
```{r}
tips$rate <- tips$tip/tips$total_bill
head(tips$rate)
average_rate = summary(tips$rate)[4]
average_rate
```

Response:The average rate is 0.1608026. This means that on an average the waiter gets $0.16 per bill.

8. Make a boxplot of the tip rate by time. The x-axis should be ordered by lunch and then dinner. Use `ggplot2` to make a chart. Verbally explain the chart.
```{r}
p <- ggplot(tips, aes(x=time, y=rate)) + 
  geom_boxplot()
p
```

Response: The tip rate during lunch is lower than during dinner.There are outliers during dinner where the trip rate is higher than normal.

9. How does smoking behavior and gender of the person who pays impact the relationship between tip and total bill? Find a visualization that incorporates all four variables. Interpret the result.
```{r}
a <- ggplot(tips, aes(x=smoker, y=rate)) + 
  geom_boxplot()
a
b <- ggplot(tips, aes(x=sex, y=rate)) + 
  geom_boxplot()
b
```

Reponse: The box plots show that the smokers pay more tips and males pay more tips


10. Use ggplot2 to find a graphical summary of the relationship between day of the week and gender of the person paying the bill. What can you say about this relationship?
```{r}
ggplot(aes(y = tip, x = day, fill = sex), data = tips) + geom_boxplot()
```

Response: Males always pay more tip then females irrespective of the day

Note: your submission is supposed to be fully reproducible, i.e. the TA and I will 'knit' your submission in RStudio. 
